---
title: "Practical Machine Learning"
author: "Olena Domanska"
date: "Monday, April 27, 2015"
output: html_document
---

## Loading and preprocessing the data

Let's first download the data for the project


```{r}
if (!file.exists("./pml-training.csv")) {
  fileUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
  download.file(fileUrl, dest = "pml-training.csv", mode = "wb")
  }
if (!file.exists("./pml-testing.csv")) {
  fileUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
  download.file(fileUrl, dest = "pml-testing.csv", mode = "wb")
  }
```

Now load the data

```{r}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

Let us separate the training data set into 2 subset train0, test0

```{r}
library(caret)
inTrain <- createDataPartition(y = training$classe, p = 0.7, list = F)
train0 <- training[inTrain, ]
test0 <- training[-inTrain, ]
```


Plot total acceleration for the subject Carlitos
```{r}
train1 <- transform(train0, classe = factor(classe))
sub1 <- subset(train1, user_name == "carlitos")
plot(sub1[,63], col= sub1$classe)
```

Try to use featurePlot to figure out some useful characteristics of variables:
```{r}
library(ggplot2)
featurePlot(x = train0[,c("total_accel_arm","total_accel_dumbbell","total_accel_forearm")], y = train0$classe, plots = "pairs")
```

We see clear separation by index, but nothing to single out the unique class.

I made a huge amount of the featurePlots but didn't identify any feature that is clearly separate classes.

Try to built prediction with linear discriminant analysis as it turned out that this method and the best method for the given problem do not differ much in accuracy. We will take that features that have not "NA" values in the testing set, that will be 51 variables 
```{r}
modlda <- train(classe ~ roll_belt+pitch_belt+yaw_belt+total_accel_belt+gyros_belt_x+gyros_belt_y+gyros_belt_z+accel_belt_x+accel_belt_y+accel_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+roll_arm+pitch_arm+yaw_arm+total_accel_arm+gyros_arm_x+gyros_arm_y+gyros_arm_z+accel_arm_x+accel_arm_y+accel_arm_z+magnet_arm_x+magnet_arm_y+magnet_arm_z+roll_dumbbell+pitch_dumbbell+yaw_dumbbell+total_accel_dumbbell+gyros_dumbbell_x+gyros_dumbbell_y+gyros_dumbbell_z+accel_dumbbell_x+accel_dumbbell_y+accel_dumbbell_z+magnet_dumbbell_x+magnet_dumbbell_y+magnet_dumbbell_z+roll_forearm+pitch_forearm+yaw_forearm+gyros_forearm_x+gyros_forearm_y+gyros_forearm_z+accel_forearm_x+accel_forearm_y+accel_forearm_z+magnet_forearm_x+magnet_forearm_y+magnet_forearm_z, method="lda", data = train0)
print(modlda)
```

Now compare the predicted values with true values:
```{r}
pred <-predict(modlda, test0)
plot(pred, unclass(test0$classe), data = test0)
table(pred, test0$classe)
```



library(knitr)
knit2html("Prediction_PML.Rmd")
browseURL("Prediction_PML.html")
